---
title: "Impute N Factors"
author: "last run by: `r Sys.info()['user']`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here("docs", "1_sanfact")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
suppressPackageStartupMessages({
  library(startR)
  library(sf)
  library(tidyverse)
})
source(here::here("code", "file_names.R"))
```

```{r}
nitrogen_factors <- read.csv(nitrogen_factors_fn,
                             stringsAsFactors = F)
```

```{r}
nitrogen_factors %>% 
  filter(rural_n > 0,
         urban_n > 0) %>% 
  select(iso3, national_n, rural_n, urban_n) %>% 
  gather(variable, value, -c(iso3, national_n)) %>% 
  ggplot(aes(x = national_n, y = value)) +
  geom_point() +
  facet_wrap(~variable) +
  coord_equal() +
  theme_bw()
```


```{r}
# Read in the world vector data, our ultimate target
world_vector_n <- st_read(world_vector_fn) %>% 
  st_set_geometry(NULL) %>% 
  select(ISO3) %>% 
  distinct() %>% 
  left_join(nitrogen_factors, by = c("ISO3" = "iso3"))

# Which ones are missing?
world_vector_n %>% 
  filter(is.na(rural_n)) %>% 
  group_by(ISO3) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(name = countrycode::countrycode(ISO3, "iso3c", "country.name"),
         region = countrycode::countrycode(ISO3, "iso3c", "region")) %>% 
  select(-n) %>% 
  knitr::kable()
```

As stated by the warning message above, the region for some of these places could not be determined. However, we can manually do it. The ones we will modify are ATF, CCK, CXR, and IOT, as seen below. In the following chunk, we create the "region" variable, and use it to group our data to calculate region-level medians for each rural / urban n factors. We then use the median to replace the missing values.

```{r}
impute_r <- function(x, medians){
  where <- is.na(x)
  x[where] <- medians[where]
  return(x)
}
```

```{r}
normalize_score <- function(var, where, norm){
  var[where] <- var[where] / norm[where]
  return(var)
}
```



```{r}
# Start imputation
nitrogen_factors_imputed <- world_vector_n %>%
  mutate(region = countrycode::countrycode(
    sourcevar = ISO3,
    origin = "iso3c",
    destination = "region"
  )) %>%
  mutate(
    region = case_when(
      ISO3 == "ATF" ~ "Eastern Africa",
      ISO3 == "CCK" ~ "South-Eastern Asia",
      ISO3 == "CXR" ~ "South-Eastern Asia",
      ISO3 == "IOT" ~ "Southern Asia",
      T ~ region
    )
  ) %>%
  select(ISO3, region, population_thousands, contains("n")) %>%
  mutate(imputed_urban_n_open = is.na(urban_n_open),
         imputed_urban_n_septic = is.na(urban_n_septic),
         imputed_urban_n_treated = is.na(urban_n_treated),
         imputed_rural_n_open = is.na(rural_n_open),
         imputed_rural_n_septic = is.na(rural_n_septic),
         imputed_rural_n_treated = is.na(rural_n_treated)) %>% 
  group_by(region) %>%                                                 # Calculate group medians before we do any imputations
  mutate(
    national_n_median = median(national_n, na.rm = T),
    rural_n_river_median = median(rural_n_river, na.rm = T),
    urban_n_river_median = median(urban_n_river, na.rm = T),
    rural_n_open_median = median(rural_n_open, na.rm = T),
    urban_n_open_median = median(urban_n_open, na.rm = T),
    rural_n_septic_median = median(rural_n_septic, na.rm = T),
    urban_n_septic_median = median(urban_n_septic, na.rm = T),
    rural_n_treated_median = median(rural_n_treated, na.rm = T),
    urban_n_treated_median = median(urban_n_treated, na.rm = T)
  ) %>%
  ungroup() %>%                                                         # If values are missing at the dev-inf, use national-inf
  mutate(
    rural_n_river = impute_r(rural_n_river, national_n_river),
    urban_n_river = impute_r(urban_n_river, national_n_river),
    rural_n_open = impute_r(rural_n_open, national_n_open),
    urban_n_open = impute_r(urban_n_open, national_n_open),
    rural_n_septic = impute_r(rural_n_septic, national_n_septic),
    urban_n_septic = impute_r(urban_n_septic, national_n_septic),
    rural_n_treated = impute_r(rural_n_treated, national_n_septic),
    urban_n_treated = impute_r(urban_n_treated, national_n_septic),
  ) %>%                                                                 # If values are still missing, use the geopolitical median
  mutate(
    national_n = impute_r(national_n, national_n_median),
    rural_n_river = impute_r(rural_n_river, rural_n_river_median),
    urban_n_river = impute_r(urban_n_river, urban_n_river_median),
    rural_n_open = impute_r(rural_n_open, rural_n_open_median),
    urban_n_open = impute_r(urban_n_open, urban_n_open_median),
    rural_n_septic = impute_r(rural_n_septic, rural_n_septic_median),
    urban_n_septic = impute_r(urban_n_septic, urban_n_septic_median),
    rural_n_treated = impute_r(rural_n_treated, rural_n_treated_median),
    urban_n_treated = impute_r(urban_n_treated, urban_n_treated_median)
  ) %>% 
  mutate(                                                         # Replace any missing values (for NFK) with the national median imputed above
    rural_n_river = impute_r(rural_n_river, national_n),
    urban_n_river = impute_r(urban_n_river, national_n),
    rural_n_open = impute_r(rural_n_open, national_n),
    urban_n_open = impute_r(urban_n_open, national_n),
    rural_n_septic = impute_r(rural_n_septic, national_n),
    urban_n_septic = impute_r(urban_n_septic, national_n),
    rural_n_treated = impute_r(rural_n_treated, national_n),
    urban_n_treated = impute_r(urban_n_treated, national_n),
  ) %>%
  select(-contains("_median"), contains("national"), -c("urban_n", "rural_n"), contains("imputed"))
```

```{r}
nitrogen_factors_imputed %>% 
  select(ISO3, national_n, contains("rural_n"), contains("urban_n"), -contains("river"), contains("imputed")) %>% 
  pivot_longer(cols = 3:14, names_to = c(".value", "set"),
               names_pattern = "(.)(.)")
```


```{r}

imputed_dummies <- nitrogen_factors_imputed %>% 
  select(ISO3, national_n, contains("imputed")) %>% 
  gather(variable, imputed, -c(ISO3, national_n)) %>% 
  mutate(variable = str_remove(variable, "imputed_"))

imputed_n_plot <- nitrogen_factors_imputed %>% 
  select(ISO3, national_n, contains("rural_n"), contains("urban_n"), -contains("river"), -contains("imputed")) %>% 
  gather(variable, value, -c(ISO3, national_n)) %>% 
  left_join(imputed_dummies, by = c("ISO3", "national_n", "variable")) %>% 
  separate(col = variable, into = c("development", "n", "treatment"), sep = "_") %>% 
  select(-n) %>% 
  ggplot(aes(x = national_n, y = value, fill = imputed)) +
  geom_point(color = "black", shape = 21) +
  facet_grid(development ~ treatment) +
  coord_equal() +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  theme(legend.justification = c(1, 1),
        legend.position = c(1, 1)) +
  labs(x = "National N factor",
       y = "N factor\n(Treatment- and development-specific )",
       fill = "Imputed")

lazy_ggsave(plot = imputed_n_plot,
            filename = "supp_n_factor_imputation",
            width = 20,
            height = 15)
```


```{r}
file.remove(nitrogen_factors_imputed_fn)
write.csv(nitrogen_factors_imputed,
          file = nitrogen_factors_imputed_fn,
          row.names = F)
```

# Reproducibility

```{r}
sessionInfo()
```

```{r}
Sys.info()
```



