---
title: "Pixel-level sanitation factors"
author: "last run by: `r Sys.info()['user']`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here("docs", "1_sanfact")) })
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(here)
  library(raster)
  library(fasterize)
  library(sf)
  library(tidyverse)
})

# set file paths
source(here("code", "file_names.R"))
```

Sanitation factors were calculated at the country level (for rural and urban) by Gordon [here](). Here, we'll take those values and assign them to 1km pixels to obtain a gridded version of country-level sanitation factors dat account for urban / rural characteristics, as well as rivers. That gridded version can then be multiplied times population density (at the pixel level) total effluent FIO effluent.

Each country has two values of sanitation factors. One value is for urban, and another value is for rural, as seen here:

```{r sanitation_factors}
sanit <- read.csv(sanit_factors_impute_fn,
                  stringsAsFactors = F) %>% 
  select(ISO3, contains("_sf"), imputed_ml, imputed_r)

sanit
```

We can look at these values on a map

```{r spatial_sf}
world_vector_sf <- st_read(world_vector_fn) %>% 
  left_join(sanit)


plot(world_vector_sf[, "urban_sf"], main = "If all was urban")
plot(world_vector_sf[, "rural_sf"], main = "If all was rural")
plot(world_vector_sf[, "imputed_ml"], main = "Imputed values")
```

Cascade has provided a raster that classifies population into urban and rural. It looks like this:

```{r rural_urban_raster}
rural_urban_input <- raster(rural_urban_binary_fn)
# Codes: -200 = NaN, 10 = Water, 111 = Rural, 222 = Urban
rec <- c(-201, 11, NA, 11, 112, 0, 113, 223, 1)
recmat <- matrix(rec, ncol = 3, byrow = T)

rural_urban_input <- reclassify(x = rural_urban_input,
                                rcl = recmat)
```

```{r river inpunt}
river <- raster(riv_15s_coastlines_fn) #load river raster
```


The goal is then to take the country-level sf in the first figure, and assign each pixel in the second with the appropriate sf value. First, let's rasterize the countries using the sf values.

## Rasterize

Now that the fudge factors are attributed to polygons, we'll rasterize them.

```{r rasterize}
sf_raster_urban <- fasterize(sf = world_vector_sf,
                             raster = rural_urban_input,
                             field = "urban_sf")

sf_raster_rural <- fasterize(sf = world_vector_sf,
                             raster = rural_urban_input,
                             field = "rural_sf")

sf_raster_urban_septic <- fasterize(sf = world_vector_sf,
                             raster = rural_urban_input,
                             field = "urban_sf_septic")

sf_raster_rural_septic <- fasterize(sf = world_vector_sf,
                             raster = rural_urban_input,
                             field = "rural_sf_septic")
```

```{r combine}
sf_raster_rural_urban <- sf_raster_urban
sf_raster_rural_urban[rural_urban_input == 0] <- sf_raster_rural[rural_urban_input == 0]
```


```{r visualize}
plot(sf_raster_rural_urban, main = "FIO values for urban / rural populations")
```

```{r combine septic}
sf_raster_rural_urban_septic <- sf_raster_urban_septic
sf_raster_rural_urban_septic[rural_urban_input == 0] <- sf_raster_rural_septic[rural_urban_input == 0]
```

```{r combine rivers}
sf_raster_rural_urban[river == 0] <- sf_raster_rural_urban_septic[river == 0]
```

```{r visualize again}
plot(sf_raster_rural_urban, main = "SF values for urban/rural populations, accounting for rivers")
```


```{r export}
file.remove(sf_raster_rural_urban_fn)
writeRaster(x = sf_raster_rural_urban,
            filename = sf_raster_rural_urban_fn)
```





















