---
title: "Pixel-level nitrogen removal factors for each infrastructure"
author: "last run by: `r Sys.info()['user']`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here("docs", "1_sanfact")) })
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(here)
  library(raster)
  library(fasterize)
  library(sf)
  library(tidyverse)
})

# set file paths
source(here("code", "file_names.R"))
```

Sanitation factors were calculated at the country level (for rural and urban) by Gordon. Here, we'll take those values and assign them to 1km pixels to obtain a gridded version of country-level sanitation factors dat account for urban / rural characteristics as well as river presence.

Each country has four values of sanitation factors. These represent the factorial combinations between urban/rural and river/no river. We do this because centers away from rivers (more than 1 km) are unlikely to deposit N into the watershed.

```{r sanitation_factors}
n_factors <- read.csv(nitrogen_factors_imputed_fn,
                     stringsAsFactors = F) %>% 
  select(ISO3, contains("_n"), -contains("national"))

n_factors
```

We can look at these values on a map

```{r spatial_sf}
world_vector_sf <- st_read(world_vector_fn) %>% 
  left_join(n_factors, by = c("ISO3" = "ISO3"))
```

Cascade has provided a raster that classifies population into urban and rural. It looks like this:

```{r rural_urban_raster}
rural_urban_input <-raster(rural_urban_binary_fn)
# Codes: -200 = NaN, 10 = Water, 111 = Rural, 222 = Urban
rec <- c(-201, 11, NA, 11, 112, 0, 113, 223, 1)
recmat <- matrix(rec, ncol = 3, byrow = T)

rural_urban_input <- reclassify(x = rural_urban_input,
                                rcl = recmat)
```

```{r river inpunt}
river <- raster(riv_15s_coastlines_fn) #load river raster
```

The goal is then to take the country-level sf in the first figure, and assign each pixel in the second with the appropriate sf value. First, let's rasterize the countries using the sf values.

## Rasterize

Now that the fudge factors are attributed to polygons, we'll rasterize them.

### Open

```{r rasterize_open}
# Infrastructure-specific
## Rural
n_raster_rural_open <- fasterize(sf = world_vector_sf,
                                 raster = rural_urban_input,
                                 field = "rural_n_open")

n_raster_urban_open <- fasterize(sf = world_vector_sf,
                                 raster = rural_urban_input,
                                 field = "urban_n_open")
```

```{r combine_open}
n_raster_rural_urban_open <- n_raster_urban_open
n_raster_rural_urban_open[rural_urban_input == 0] <- n_raster_rural_open[rural_urban_input == 0]
n_raster_rural_urban_open <- river * n_raster_rural_urban_open
```

```{r}
file.remove(nitrogen_factors_open_raster_fn)
writeRaster(x = n_raster_rural_urban_open,
            filename = nitrogen_factors_open_raster_fn)

# Remove rasters to save space
rm(n_raster_urban_open)
rm(n_raster_rural_open)
```


### Septic

```{r rasterize_septic}
# Infrastructure-specific
n_raster_rural_septic <- fasterize(sf = world_vector_sf,
                            raster = rural_urban_input,
                            field = "rural_n_septic")

n_raster_urban_septic <- fasterize(sf = world_vector_sf,
                            raster = rural_urban_input,
                            field = "urban_n_septic")
```

```{r combine_septic}
n_raster_rural_urban_septic <- n_raster_urban_septic
n_raster_rural_urban_septic[rural_urban_input == 0] <- n_raster_rural_septic[rural_urban_input == 0]
n_raster_rural_urban_septic <- river * n_raster_rural_urban_septic
```

```{r export_septic}
file.remove(nitrogen_factors_septic_raster_fn)
writeRaster(x = n_raster_rural_urban_septic,
            filename = nitrogen_factors_septic_raster_fn)

# Remove rasters to save space
rm(n_raster_urban_septic)
rm(n_raster_rural_septic)
```

```{r visualize_septic}
rasterVis::levelplot(n_raster_rural_urban_septic)
```


### Treated

```{r rasterize_treated}
n_raster_rural_treated <- fasterize(sf = world_vector_sf,
                            raster = rural_urban_input,
                            field = "rural_n_treated")

n_raster_urban_treated <- fasterize(sf = world_vector_sf,
                            raster = rural_urban_input,
                            field = "urban_n_treated")
```

```{r combine_treated}
n_raster_rural_urban_treated <- n_raster_urban_treated
n_raster_rural_urban_treated[rural_urban_input == 0] <- n_raster_rural_treated[rural_urban_input == 0]
```

```{r export_treated}
file.remove(nitrogen_factors_treated_raster_fn)
writeRaster(x = n_raster_rural_urban_treated,
            filename = nitrogen_factors_treated_raster_fn)

# Remove rasters to save space
rm(n_raster_urban_treated)
rm(n_raster_rural_treated)
```


```{r visualize}
rasterVis::levelplot(n_raster_rural_urban_treated)
```























