---
title: "Pour Point Processing"
author: "Jared Kibele"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    pandoc_args: ["--output=../docs/pour_points.html"]
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)

source("./file_names.R")
```

## Blank File Prep

The original file came from a previous analysis (as described [here](datasets.html)). It's in a Mollweide projection. I'm still working in unprojected WGS84 (to match the population raster), so I'll first reproject. Then I'll drop the fields I don't need and add a field for my effuluent density sums (initialized with zeros).

```{r file_prep}
drop_cols <- c('SUM_FERTC','SUM_PESTC', 'SUM_IMPV')

glob_plume <- sf::read_sf(pour_points_original_fn) %>% 
  sf::st_transform(4326) %>% 
  select(-one_of(drop_cols)) 
```

## Join Watershed Effluent Values to Pour Points

```{r watershed_join}
watersheds <- sf::read_sf(effluent_watersheds_fn) %>% 
  as_tibble() %>% 
  select(c("basin_id", "effluent", "count", "area"))

pour_points <- glob_plume %>% 
  left_join(watersheds, by="basin_id") %>% 
  filter(effluent > 0.0)

pour_points %>% 
  sf::write_sf(pour_points_fn,
               layer_options = "OVERWRITE=true")

pour_points %>% head()
```

Writing the shapefile produces an error message, but if I'm reading [this](https://github.com/r-spatial/sf/issues/306) correctly we're not actually losing any data. TODO: Double check that we can disregard.

```{r eval=FALSE}
# This takes forever. I don't have the patients right now so `eval=FALSE`
ggplot(pour_points) +
  geom_sf(aes(size=effluent))
```