---
title: "Sanitation Factors"
author: "Jared Kibele"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    pandoc_args: ["--output=../docs/sanitation_factors.html"]
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE}
library(tidyverse)

# set file paths
source("./file_names.R")
```


We need to map the 5 sanitation categories from the WHO/Unicef [data](https://github.com/jkibele/wastewater/issues/3#issuecomment-460335655) to a set of "sanitation factors" based on [Kay et al., 2008](https://github.com/jkibele/wastewater/issues/3#issue-402495556). I need to do more reading, but I'm thinking the sanitation factors ($sf_{category}$) will range from 1.0 (no reduction in excreta) for "open defication" to (just a random placeholder guess) 0.3 (70% reduction) for "Safely Managed". I need to go back and read Kay et al. with this idea in mind to pick a reasonalbe set of sanitation factors. Once we have the set of sanitation factors, we can use the sanitation category proportions to come up with a national weighted average sanitation factor for each country ($sf_{country}$). 


## 1) Developing National Sanitation Factors

Map santiation categories to $sf_{category}$. As a first pass, I'm only going to deal with the national figures that do not separate rural and urban. ~~If we had a global rural/urban dataset (maybe based on land cover), we could refine this a bit~~. (too much missing data for that) But, for now, I'm just going to read the part of the spreadsheet that has the national level data. I'm also going to just focus on 2015 values (that's what the [pop density data](https://github.com/jkibele/wastewater/issues/1#issuecomment-459908336) is adjusted to reflect.)

```{r sf_mapping, cache=FALSE, include=TRUE, warning=FALSE, message=FALSE}

# Read the relevant parts of the spreadsheet
sanit_all <- readxl::read_excel(san_dat, 
                            sheet="Sanitation",
                            range=readxl::cell_limits(c(3,1), c(NA,33)),
                            na = "-"
                            )
# Prepare to rename the sanitation categories with coding friendly
# single word versions of the categories
nm_old <- colnames(sanit_all)[6:9]
nm_new <- c(
  "Basic",
  "Limited",
  "Unimproved",
  "Open"
)

# Code in sanitation factors for categories (just making these up for now)
sf_cat <- list(
  Basic = 0.3,
  Limited = 0.5,
  Unimproved = 0.8,
  Open = 1.0
)

# Filter to just 2015 and rename categories
sanit <- sanit_all %>% 
  filter(Year == 2015) %>% 
  rename_at(vars(nm_old), ~ nm_new) %>% 
  mutate(sf = Basic * 0.01 * sf_cat[['Basic']] +
              Limited * 0.01 * sf_cat[['Limited']] +
              Unimproved * 0.01 * sf_cat[['Unimproved']] +
              Open * 0.01 * sf_cat[['Open']]
           ) %>% 
  select(-c(10:26, "% urban"))

sanit %>% write_csv(
  sanit_factors_fn,
  col_names = TRUE
)
```

Here are what the sanitation data look like ($sf$ is at the far right):

```{r sf_disp, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
options(knitr.kable.NA = '--')
# Just display a table to show where we're at
sanit %>% 
  kableExtra::kable(digits = 2) %>% 
  kableExtra::add_header_above(c(" " = 4, "National" = 4, "Breakdown of Basic" = 7, " " = 1)) %>% 
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover")
    ) %>% 
  kableExtra::scroll_box(width = "100%", height = "400px") 
```


## To Do

The current sanitation factors are just place holder. We need to refine the formula for calculating them. For now, the intermediate output can be found at ``r sanit_factors_fn``.
