---
title: "Population Spatial Overlap"
author: "Jared Kibele"
date: "3/13/2019"
output: 
  html_document: 
    pandoc_args: ["--output=../docs/spatial_overlap.html"]
---

I need to figure out which country dataset is better for joining up with sanitation data and the population data. The contestants are `regions_2017_update` and `gadm`. I'm going to check the spatial overlap with population data here. Here's how I'm going to do it:

1. Starting with the population raster, I'll make a raster with the value 2 in every pixel where population density is > 0 and the value 0 elsewhere.
2. I'll make a raster from each of my vector layers (GADM and Regions) with 1s where there's a polygon and 0s where there's not.
3. Then I'll add the pop based raster to each of the vector derived raster. Values of 3 will represent overlap, 1 will represent vector areas with no population, and 2s will represent population data without overlapping vector (in other words, population data that we can't link to sanitation data).

```{r inspect_pop_raster}
library(tidyverse)
data_dir <- "/home/shares/ohi/git-annex/land-based/wastewater"
pop_fn <- file.path(data_dir, "d2019/gpw-v4-population-density-adjusted-to-2015-unwpp-country-totals-rev10_2015_30_sec_tif/gpw_v4_population_density_adjusted_to_2015_unwpp_country_totals_rev10_2015_30_sec.tif")
reg_fn <- file.path(data_dir, "regions_2017_update/regions_2017_update.shp")
gadm_fn <-  file.path(data_dir, "gadm36_shp/gadm36.shp")

# For later use in bash chunks
Sys.setenv(POP_FN=pop_fn, REG_FN=reg_fn, GADM_FN=gadm_fn)

# Check the resolution of the population raster
cmd <- paste("gdalinfo", pop_fn, "| grep -E 'Pixel Size|Upper|Lower'")
system(cmd)
```

```{r build_rasters, cache=TRUE}

# Make file paths for our outputs
pop2_fn <- file.path(data_dir, "pop2.tif")
reg_exists_fn <- file.path(data_dir, "reg_exists.tif")
gadm_exists_fn <- file.path(data_dir, "gadm_exists.tif")
reg_calc_fn <- file.path(data_dir, "reg_overlap.tif")
gadm_calc_fn <- file.path(data_dir, "gadm_overlap.tif")

# Set env for later use in bash chunk
Sys.setenv(POP_FN=pop_fn,
           POP2_FN=pop2_fn, 
           REG_EXISTS_FN=reg_exists_fn, 
           GADM_EXISTS_FN=gadm_exists_fn,
           REG_CALC_FN=reg_calc_fn,
           GADM_CALC_FN=gadm_calc_fn)

# GDAL commands are weird about overwriting, so delete files first
# file.remove(pop2_fn)
# 
# pop2_cmd <- sprintf('gdal_calc.py -A %s --outfile=%s --type=Byte --calc="2*(A>0)"', pop_fn, pop2_fn)
# system(pop2_cmd)

```

```{bash}
gdal_calc.py --overwrite --NoDataValue=0 -A $POP_FN --outfile=pop2_temp.tif --type=Byte --calc="2*(A>0)"
gdal_translate -a_nodata none -ot Byte pop2_temp.tif $POP2_FN
rm pop2_temp.tif
```


Check the projection of the regions shapefile:

```{bash}
ogrinfo -al -so $REG_FN
```
 
 I want to keep everything in WGS84 (the projection of the population data) for now. So I'll use OGR to reproject. Then I'll rasterize the reprojected
 
```{bash}
ogr2ogr -overwrite -t_srs EPSG:4326 regions_4326.shp $REG_FN # destination before source, for some reason
rm $REG_EXISTS_FN # Because rasterize doesn't have an overwrite option
gdal_rasterize -init 0 -burn 1 -tr 0.008333333333334 0.008333333333334 \
-te -180 -60 180 85 -ot Byte regions_4326.shp $REG_EXISTS_FN
#rm regions_4326.shp
```
 
 Now do the band calc.
 
```{bash}
gdal_calc.py -A $REG_EXISTS_FN -B $POP2_FN --outfile=$REG_CALC_FN --type=Byte --calc="A+B"
```
```{bash}
rm $GADM_EXISTS_FN
gdal_rasterize -init 0 -burn 1 -tr 0.008333333333334 0.008333333333334 \
-te -180 -60 180 85 -ot Byte $GADM_FN $GADM_EXISTS_FN
```
```{bash}
gdal_calc.py -A $GADM_EXISTS_FN -B $POP2_FN --outfile=$GADM_CALC_FN --type=Byte --calc="A+B"
```
 
```{r}
library(raster)
library(leaflet)
pnw_extent <- extent(-127, -120, 46.8, 51)
gadm_calc <- raster(gadm_calc_fn)
gadm_crop <- gadm_calc %>% crop(pnw_extent, snap='in')
```

```{r}
color_labels <- c("No pop, no vector",
                  "Vector, no pop",
                  "Pop, no vector",
                  "Overlap")
leaflet() %>% 
  addTiles() %>%
  addRasterImage(gadm_crop %>% as.factor(), 
                 colors=c("#86b1c8", "#0039d6", "#be0026", "#04b038")) %>% 
  addLegend(colors=c("#86b1c8", "#0039d6", "#be0026", "#04b038"),
            labels = color_labels)
```



