---
title: "Pixel-level effluent"
author: "last run by: `r Sys.info()['user']`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here("docs")) })
---

```{r setup, include = FALSE, message = FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(here)
  library(raster)
  library(tidyverse)
})

# set file paths
source(here("code", "file_names.R"))
```

Here I will calculate pixel-level effluent. This is the product of population size in each pixel, the country-level per capita yearly caloric intake, and the sanitation factors derived before. Therefore:

$$
E_i = P_i \times SF_i
$$

First, I'll read the three rasters.

```{r load data}
pop <- raster(pop_fn)
sf_raster_rural_urban <- raster(sf_raster_rural_urban_fn)
```

Then, multiply one times the other.

```{r multiply rasters}
effluent <- pop * sf_raster_rural_urban
```

Then, we visualize the "effluent". The visualization is in log-scale, but  the data are not really in log-scale.

```{r plot preview}
plot(calc(effluent, log), main = "log-transformed effluent (pop * sf_low)")
```

And then we export it as a tiff

```{r export}
file.remove(effluent_FIO_fn)
writeRaster(x = effluent,
            filename = effluent_FIO_fn)
```

