---
title: "gap_filling"
author: "Gordon Blasco"
date: "10/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Gap filling
```{r, warning =FALSE, message=FALSE}


library(tidyverse)
library(here)
library(countrycode)
library(sf)
library(broom)
library(WDI)
library(caret)


source(here("code", "file_names.R"))


all_missings_list <- read_csv(here("scratch","gordon","missing_data.csv"))

country_codes <- countrycode::codelist %>% 
  dplyr::select(country.name.en, iso3c,
                iso3n, iso2c 
                ) %>% 
  rename(ISO3 = iso3c)

sanit <- read.csv(sanit_factors_fn,
                  stringsAsFactors = F) %>% 
  dplyr::select(iso3, rural_sf, urban_sf)

world_vector_sf <- st_read(world_vector_fn) %>% 
  left_join(sanit, by = c("ISO3" = "iso3"))


WDI_list <- WDIcache()
wdi_names <- WDI_list[["series"]]

```



## load in gdp world bank country indicators
```{r message=FALSE}
wdi_indicators <- WDI(indicator=c('NY.GDP.PCAP.KD', '1.1_ACCESS.ELECTRICITY.TOT'), country= "all", start=2015, end=2015) %>% 
  rename(gdp_per_cap = NY.GDP.PCAP.KD,
         access_electricity = `1.1_ACCESS.ELECTRICITY.TOT`) %>% #ended up not using this
  left_join(country_codes) %>% 
  dplyr::select(-year, -country, -iso2c, -country.name.en, -iso3n) %>% 
  rename(iso3 = ISO3)

missings <- as.data.frame(world_vector_sf) %>% 
  dplyr::select(ISO3, rural_sf, urban_sf)  %>% 
  left_join(country_codes) %>% 
  distinct(ISO3, .keep_all = TRUE) %>% 
  left_join(wdi_indicators, by = c("ISO3" = "iso3"))

ggplot(missings, aes(x=log(gdp_per_cap), y = rural_sf))+
  geom_point()


```

## Build a training set with all data that does exist.
```{r message=FALSE, warning = FALSE}
# only select national level data and gdp
vars <- c(1:9, 43, 45:46)

sanfac_data <- read_csv(file.path(intermediate_dir, "clean_sanfac_2019.csv")) %>% 
  mutate_at(vars(6:44), parse_number) %>% 
  filter(year == 2015) %>% 
  dplyr::select(-sl, -year, -country) %>% 
  left_join(wdi_indicators) %>% 
  left_join(sanit) %>% 
  dplyr::select(vars)


sanfac_train <- sanfac_data %>% 
  filter(!is.na(rural_sf)) %>% 
  dplyr::select(-rural_sf, -urban_sf, -iso3)


rural_train <- sanfac_data %>% 
  filter(!is.na(rural_sf)) %>% 
  dplyr::pull(rural_sf)



```


# set the seed and run a lm 
```{r warning = FALSE}


set.seed(310)
ctrl <- trainControl(method = "repeatedcv", repeats = 5)

ruralLm <- caret::train(x = sanfac_train, y = rural_train, method = "lm")
rural_lm_info <- ruralLm$results
```


# Run a PCR and PLS regression model and compare the RMSE's with different number of components. 
```{r warning = FALSE}
set.seed(310)
ruralPCR <- caret::train(x = sanfac_train, y = rural_train,
                        method = "pcr",
                        #trControl = ctrl,
                        tuneLength =8)

set.seed(310)
ruralPLS <- caret::train(x = sanfac_train, y = rural_train,
                        method = "pls",
                        #trControl = ctrl,
                        preProcess = c("center", "scale"),
                        tuneLength =8)


comps <- rbind(ruralPLS$results, ruralPCR$results)
comps$Model <- rep(c("PLS", "PCR"), each =8)
```


## Plot the RMSE as a function of the number of components for the PCR and PLS models.
```{r}
ggplot(comps, aes(x=ncomp, y = RMSE))+
  geom_line(aes(color = Model))+
  labs(
    title = "PCR vs PLS regression for rural sanitation factors",
    subtitle = "the black line is the RMSE for a lm",
    x = "Number of Components"
  ) +
  theme_bw()+
  scale_y_continuous(limits = c(0, .25), expand = c(0,0))+
  scale_x_continuous(expand = c(0,0)) +
  geom_hline(yintercept = rural_lm_info$RMSE[1])

```

both the pcr and pls are preforming worse or as well as the the linear model. Probably best to use the lm. 

# use the lm to fill in missing data
```{r}
sanfac_fill <- sanfac_data %>% 
  filter(is.na(rural_sf)) %>% 
  dplyr::select(-rural_sf, -urban_sf) %>% 
  distinct(iso3, .keep_all = TRUE) %>% 
  column_to_rownames(var = "iso3")

test <- predict(ruralLm, newdata = sanfac_fill)

filled <- sanfac_fill %>% 
  mutate(predicted_rural_sf = test) 
```

