---
title: "Pixel-level and infrastructure-specific Nitrogen"
author: "last run by: `r Sys.info()['user']`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here("docs")) })
---

```{r setup, include = FALSE, message = FALSE, warning=FALSE}
suppressPackageStartupMessages({
  library(here)
  library(raster)
  library(tidyverse)
})

# set file paths
source(here("code", "file_names.R"))
```

Here I will calculate pixel-level effluent for each type of infrastructure (open, septic, treated). This is the product of population size in each pixel, the country-level per capita yearly protein intake, and the sanitation factors derived before. Therefore:

$$
N_i = P_i \times NR_i \times 0.16 \times p_i
$$

In this case, $Nr$ will be the infrastructure-specific values.

First, I'll read the three rasters.

```{r load data}
pop <- raster(pop_fn)
nitrogen_factors_open_raster <- raster(nitrogen_factors_open_raster_fn)
nitrogen_factors_septic_raster <- raster(nitrogen_factors_septic_raster_fn)
nitrogen_factors_treated_raster <- raster(nitrogen_factors_treated_raster_fn)
protein <- raster(protein_g_cap_year_raster_fn)
```

Then, multiply one times the other.

```{r multiply rasters}
n_open <- pop * nitrogen_factors_open_raster * protein * 0.16
n_septic <- pop * nitrogen_factors_septic_raster * protein * 0.16
n_treated <- pop * nitrogen_factors_treated_raster * protein * 0.16
```

```{r calculate_total_n}
n_total <- n_open + n_septic + n_treated
n_total_log10 <- calc(n_total, log10)
```


Remove files to save RAM

```{r remove_variables}
rm(pop)
rm(protein)
rm(nitrogen_factors_open_raster)
rm(nitrogen_factors_septic_raster)
rm(nitrogen_factors_treated_raster)
```


Then, we visualize the "effluent". The visualization is in log-scale, but  the data are not really in log-scale.

```{r plot preview}
effluent_N_total_log10 <- calc(n_total, log10)
rasterVis::levelplot(effluent_N_total_log10)
```

```{r log_10_transform_attributions}
n_open_log10 <- calc(n_open, log10)
n_septic_log10 <- calc(n_septic, log10)
n_treated_log10 <- calc(n_treated, log10)
```


And then we export them all as tiff files

```{r export_total}
file.remove(effluent_N_fn)
writeRaster(x = n_total,
            filename = effluent_N_fn)
```

```{r export_total_log10}
file.remove(effluent_N_log10_fn)
writeRaster(x = effluent_N_total_log10,
            filename = effluent_N_log10_fn)
```


```{r export_open}
file.remove(effluent_N_open_fn)
writeRaster(x = n_open,
            filename = effluent_N_open_fn)
```


```{r export_septic}
file.remove(effluent_N_septic_fn)
writeRaster(x = n_septic,
            filename = effluent_N_septic_fn)
```


```{r export_treated}
file.remove(effluent_N_treated_fn)
writeRaster(x = n_treated,
            filename = effluent_N_treated_fn)
```

```{r export_open_log10}
file.remove(effluent_N_open_log10_fn)
writeRaster(x = n_open_log10,
            filename = effluent_N_open_log10_fn)
```


```{r export_septic_log10}
file.remove(effluent_N_septic_fn_log10)
writeRaster(x = n_septic_log10,
            filename = effluent_N_septic_log10_fn)
```


```{r export_treated_log10}
file.remove(effluent_N_treated_log10_fn)
writeRaster(x = n_treated_log10,
            filename = effluent_N_treated_log10_fn)
```


